---
title: "Exercise 01 - dplyr -- data wrangling"

date-modified: 'today'
date-format: long

format: 
  html:
    footer: "CC BY 4.0 John R Little"

license: CC BY    
---

## Goals:

Use dplyr functions to wrangle data

In this exercise you'll practice the `dplyr` verbs (i.e. *functions*) mentioned in the demonstration.

-   `arrange()`
-   `filter()`
-   `mutate()`
-   `select()`
-   `count()`

## Directions

### Setup

We've prebuilt a data set that has information about menus in Duke's Brodhead Center (dining hall) from three restaurants. Create a new R Notebook, load the *tidyverse* package, and **load the dining hall dataset**, *as a tibble*. The broadhead dataset is a CSV file in the `data` directory.

a.  Add a code chunk and Load the `tidyverse` library package

```{r}
#| warning: false
#| message: false
library(tidyverse)
```

b.  Insert a code chunk and import (`read_csv()`) the brodhead practice data from the `data` directory. Assign the imported data to an object name, e.g. "brodhead"

```{r}
#| warning: false
#| message: false
brodhead <- read_csv("../data/brodhead_center.csv")
```

Now you are ready to make new code chunks and follow the steps in Exercise 1 (below).

> Answers can be found in the exercise_01_answers.Rmd file

### Exercise: Data Wrangling

All of the following questions are based on the sample of restaurants represented in the dataset which is accurate as of September 2, 2016.

1.  Which restaurant has the lowest cost item and what is the item?

```{r}
brodhead %>% 
  _______(cost) %>% 
  select(name, type, itemName, cost)
```

2.  Which restaurant has the most expensive item(s)? What are those item(s)?

```{r}
brodhead %>% 
  ____(____(cost)) %>% 
  select(name, type, itemName, cost)
```

3.  At the Brodhead Center, how many of the entrees (found in the `menuType` variable) cost eight dollars?

```{r}
brodhead %>% 
  ______(cost == _, menuType == "______") %>% 
  select(name, menuType, itemName, cost)
```

4.  The head of Duke dining is considering reducing prices at the Brodhead Center. Using what we've learned in class, write code that will calculate a new variable (`halfPrice`) that contains items at half price.

```{r}
brodhead %>% 
  ______(halfPrice = (____ / 2)) %>% 
  select(name, itemName, cost, halfPrice)
```

5.  How many entrees are in the dataset (`menuType` variable)? How many desserts

```{r}
brodhead %>% 
  _____(menuType)
# You can use `filter()` to limit by menuType
```

## Answers

> Answers can be found in the `_exercise_01/answers/01_dplyr_answers.qmd` file

## Optional

[Interactive quiz](https://posit.cloud/learn/primers/2.2)
