---
title: "Answers - Excercise 00"

date-modified: 'today'
date-format: long

format: 
  html:
    footer: "CC BY 4.0 John R Little"

license: CC BY    
---

## exercise 00:

1.  Load the tidyverse libary package

```{r}
#| message: false
#| warning: false
library(tidyverse)
```

2.  Fill in the blanks. Using what you've seen in class, if `data/brodhead_center.csv` is a CSV (comma separated values) file, how would you load the file into a new object called `brodhead`?

```{r}
#| message: false
#| warning: false
brodhead  <- read_csv("../../data/brodhead_center.csv")

# or
# brodhead <- read_csv(here::here("data", "brodhead_center.csv"))
```

3.  Display your new `brodhead` data frame?

```{r}
brodhead
```

4.  `starwars` is an on-board dataset that comes with the tidyverse. Insert a new code chunk and display that data.

```{r}
starwars
```

5.  Copy and paste the following code into a new code chunk in your new file.

```{r}
starwars %>% 
  ggplot(aes(fct_infreq(hair_color))) + 
  geom_bar() + 
  ggtitle("Hair Color Frequency of Star Wars Characters")
```

6.  Take a look at the structure of the `brodhead` object.
    -   How many observations (rows) are there?
    -   How many variables (columns) are there?
    -   How many of the variables are numeric data?

```{r}
glimpse(brodhead)
```

## Bonus 2

Importing labeled vectors from an spss file format. Import via the `haven::read_sav()` function

```{r}
library(haven)
library(here)

my_labeled_vectors_df <- 
  read_sav(here("data", "student_satisfaction_test-data_from_qualtrics.sav"))

my_labeled_vectors_df

my_labeled_vectors_df |> 
  select(Q2) |> 
  mutate(Q2_labels = as_factor(Q2))

my_labeled_vectors_df |> 
  select(ResponseId, starts_with("Q")) |> 
  mutate(across(is.labelled, function(x) as_factor(x), .names = "{.col}_label")) |>
  select(ResponseId, sort(tidyselect::peek_vars()))

```
