---
title: "Answers 01 - dplyr -- data wrangling"

date-modified: 'today'
date-format: long

format: 
  html:
    footer: "CC BY 4.0 John R Little"

license: CC BY
---

### Setup

a.  Load the `tidyverse` library package

```{r}
#| message: false
#| warning: false
library(tidyverse)
```

b.  Import data

```{r}
#| message: false
#| warning: false
library(here)
brodhead <- read_csv(here("data", "brodhead_center.csv"))
```

### exercise_01 -- Data Wrangling

1.  Which restaurant has the lowest cost item and what is the item?

```{r}
brodhead %>% 
  arrange(cost) %>% 
  select(name, type, itemName, cost)
```

#### Alternative Answer

```{r}
lowest_cost_item <- brodhead %>% 
  arrange(cost) %>% 
  select(name) %>% 
  slice_head() %>% 
  pull()
lowest_cost_item  
```

The lowest cost item can be found at `r lowest_cost_item`.

2.  Which restaurant has the most expensive item(s)? What are those item(s)?

```{r}
brodhead %>% 
  arrange(desc(cost)) %>% 
  select(name, type, itemName, cost)
```

#### Alternative Answers

```{r}
answer2 <- brodhead %>% 
  select(name, type, itemName, cost) %>% 
  filter(type == "restaurant",
         cost == max(cost)) 
answer2
```

```{r}
expensive_items <- brodhead %>% 
  select(name, type, itemName, cost) %>% 
  filter(type == "restaurant",
         cost == max(cost)) %>% 
  select(name, itemName) 

expensive_items
```

**Example of inline coding** (Render report to see result.)

The `r unique(expensive_items$name)` restaurant offers the following items at the highest cost: `r expensive_items$itemName`.

3.  At the Brodhead Center, how many of the entrees (found in the `menuType` variable) cost eight dollars?

```{r}
brodhead %>% 
  filter(cost == 8, menuType == "entree") %>% 
  select(name, menuType, itemName, cost)
```

The **answer** is 8. You know this because the Tibble (data frame) has 8 rows.

#### Alternative Answer

```{r}
brodhead %>% 
  filter(cost == 8, menuType == "entree") %>% 
  select(name, menuType, itemName, cost) %>% 
  count(menuType)
```

#### Alternative Answer

```{r}
eight_dollar_items <- brodhead %>% 
  filter(cost == 8, menuType == "entree") %>% 
  select(name, menuType, itemName, cost) 

length(eight_dollar_items$itemName)
```

4.  The head of Duke dining is considering reducing prices at the Brodhead Center. Using what we've learned in class, write code that will calculate a new variable (`halfPrice`) that contains items at half price.

```{r}
brodhead %>% 
  mutate(halfPrice = (cost / 2)) %>% 
  select(name, itemName, cost, halfPrice) %>% 
  mutate(halfPrice = scales::dollar(halfPrice))
```

5.  How many entrees are in the dataset (`menuType` variable)? How many desserts

```{r}
brodhead %>% 
  count(menuType)
# You can use `filter()` to limit by menuType
```

#### Alternative Answer

```{r}
brodhead %>% 
  count(menuType) %>% 
  filter(menuType == "entree") %>% 
  pull(n)
```

```{r}
#| echo: false
my_answer <- brodhead %>% 
  count(menuType) %>% 
  filter(menuType == "entree") %>% 
  pull(n)
```

There are `r my_answer` entrees.
