---
title: "03 pivot data"
author: "John Little"

date-modified: 'today'
date-format: long

format: 
  html:
    footer: "CC BY 4.0 John R Little"

license: CC BY    
---

```{r}
library(tidyverse)
```

## data

These exercises use the following [`ggplot2` training datasets](https://ggplot2.tidyverse.org/reference/index.html#section-data)

-   ggplot2::economics

## Pivot

Below are two data frames. One is wide data, the other is long.

```{r}
economics
economics_long %>% arrange(date)
```

## Goal

Using one of the dplyr pivot functions, pivot the economics data to long format

```{r}
economics %>% 
  pivot_longer(cols = pce:unemploy, 
               names_to = "variable", 
               values_to = "value") 
```

or

```{r}
economics |> 
  pivot_longer(-date, names_to = "variable", values_to = "value")
```

Now that the data are long. Can you use the `facet_wrap()` function to make multiple line plots, one line plot for each `variable` category?

```{r}
economics |> 
  pivot_longer(-date, names_to = "variable", values_to = "value") |> 
  ggplot(aes(date, value)) +
  geom_line() + 
  facet_wrap(vars(variable))
```
