---
title: "Visualization"
#output: html_document
#  html_notebook:
#    css: duke.css
---

When in the data exploratory phase, you can quickly generate very basic visualizations with minimal effort.  Your visualizations then appear in the Viewer Pane.  

For this introductory section on visualization we use the `ggvis` package because it works seamlessly with pipes (i.e. `%>%`).  You can do a lot with [ggvis](http://ggvis.rstudio.com/).  You can do even more with [ggplot2](http://ggplot2.tidyverse.org/).  We offer a separate workshop on ggplot2.

```{r loadPackages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggvis)
#library(datasets)
#library(forcats)
```


```{r not_jabba}
not_jabba <- starwars %>% 
  filter(!is.na(mass),
         !is.na(height),
         mass < 400)
```


## Scatter Plot
Linear model and 95% confidence interval

```{r ggvis-scatter}
not_jabba %>% 
  ggvis(~height, ~mass) %>% 
  layer_points() %>% 
  layer_model_predictions(model = "lm", se = TRUE)

```
[More information on ggvis scatterplots](http://ggvis.rstudio.com/cookbook.html#scatterplots) | [line graphs](http://ggvis.rstudio.com/cookbook.html#line-graphs)

## Bar Graph

```{r ggvis_prepare_bargraph}

# unique(starwars$species)

levels <- starwars %>%
  group_by(species) %>% 
  count() %>% 
  arrange(-n) %>% 
  .[1]

#glimpse(foo)

#length(foo$species)

starwars_species_as_fctr  <-  starwars %>% 
  mutate(species = parse_factor(species, levels$species))


glimpse(starwars_species_as_fctr)
glimpse(starwars$species)


starwars_species_as_fctr_foo  <-  starwars %>% 
  mutate(species2 = parse_factor(species, species))

starwars_species_as_fctr_foo %>% 
  select(species, species2, name)

levels(starwars_species_as_fctr$species)
```



starwars_levels <- starwars %>% 
  filter(!is.na(species)) %>% 
  group_by(species) %>% 
  count(species) %>% 
  arrange(-n) %>% 
  filter(n > 1) %>% 
  .[,1] %>% 
  ungroup() %>% 
  as_vector()

starwars_levels <- parse_factor(starwars_levels, starwars_levels)

class(starwars_levels)
glimpse(starwars_levels)
starwars_levels


starwars2  <-  starwars %>% 
  mutate(species = parse_factor(species, species))

glimpse(starwars2)
glimpse(starwars)



```{r ggvis_bargraph}
fooo %>% 
  mutate(species = parse_factor(species, species_levels)) %>% 
  
  
  

starwars_species_as_fctr %>%   
  ggvis(~species) %>% 
  layer_bars()
```
[More information on ggvis bar graphs](http://ggvis.rstudio.com/cookbook.html#bar-graphs)

## Histogram

```{r ggvis-hist, message=FALSE}
starwars %>% 
  ggvis(~birth_year) %>% 
  layer_histograms()
```
[More information on ggvis histograms](http://ggvis.rstudio.com/cookbook.html#histograms)

## Grouping data points

```{r ggvis-line, warning=FALSE}
starwars %>% 
  filter(!is.na(mass),
         !is.na(height),
         mass < 400,
         height > 100,
         species == "Human" | 
           species == "Droid" | 
           species == "Gungan") %>% 
  #mutate(species = parse_factor(species, species)) %>% 
  ggvis(~height, ~mass) %>% 
  group_by(species) %>% 
  #layer_points(~mass, ~height,
  layer_points(~height, ~mass, 
               stroke := "black", 
               fill = ~species)

```



```{r}
mtcars %>% 
  ggvis(~wt, ~mpg) %>% 
  group_by(cyl) %>% 
  layer_model_predictions(model="lm", se = TRUE) %>% 
  layer_points(~wt, ~mpg, stroke := "black", fill = ~factor(cyl), size := ~hp, opacity := 0.7) 
```

&nbsp;

## BoxPlot

```{r}
mtcars %>% 
  ggvis(~factor(cyl), ~mpg) %>% 
  layer_boxplots()
  
  
```

```{r, message=FALSE, warning=FALSE}
starwars %>% 
  filter(!is.na(mass),
         !is.na(height),
         species == "Human" | 
           species == "Droid" | 
           species == "Gungan") %>% 
  mutate(species = parse_factor(species, species)) %>% 
  ggvis(~parse_factor(species, species), ~mass) %>% 
  layer_boxplots()
  
```


[More information on box plots](http://ggvis.rstudio.com/cookbook.html)
