---
title: "Data Wrangling"
output: html_document
---

**dplyr** is a widely adopted r package, part of tidyverse, which uses **English verbs** to reshape data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Arrange


**Arrange** to sort rows (observations) by column headers (variables)

```{r makesmall_sw_dataframe, message=FALSE, warning=FALSE, include=FALSE}

small_sw <- starwars %>% 
  select(species, mass, homeworld, name)
```

```{r}
small_sw
```



```{r arrange, message=FALSE, warning=FALSE}
small_sw %>% 
  arrange(species, desc(mass), homeworld)
```

## Select

**Select** to choose columns (variables)

```{r select}
starwars %>% 
  arrange(species, desc(mass), homeworld) %>% 
  select(species, mass, homeworld, name:starships)
```


## Filter

**Filter** to select rows (observations)

```{r filter}
starwars_32_mass <- starwars %>% 
  filter(mass == 80)

starwars_32_mass    # display filtered data frame

```

## Mutate

**Mutate** to create new variables (columns) 

```{r mutate}
starwars_32_mass %>% 
  mutate(BMI = height / mass) %>% 
  select(name, BMI, height, mass, hair_color, 1:13)
```

## Count

**Count** to summarize observations (rows)

```{r count}
starwars %>% 
  count(mass)
```

## Summarize
Summarize to collapse values into a summary

```{r}
starwars %>% 
  group_by(species) %>% 
  summarize(Count = n(), 
            mean_ht = mean(height), 
            min_ht = min(height), 
            max_ht = max(height)) %>% 
  arrange(desc(Count))
```

```{r include=FALSE}
ungroup(starwars)
```


## Put it all together

We can pipe commands together.  Think of a pipe (i.e. `%>%`) as a conjunction.  Any time you see a pipe, think "and then."  You can insert a pipe with 'Ctrl+Shift+M' (Help > Keyboard Shortcts)

```{r alltogether}
starwars %>% 
  group_by(species) %>% 
  summarize(Count = n(), 
            mean_ht = mean(height), 
            min_ht = min(height), 
            max_ht = max(height)) %>% 
  arrange(desc(Count)) %>% 
  filter(!is.na(mean_ht), Count > 1)
```
