---
title: "Functions in R"

date-modified: 'today'
date-format: long

license: CC BY-NC
bibliography: references.bib
---

R is a functional programming language.[^1] This means repetitive execution is emphasized without using *FOR loops.*[^2] *No FOR loops*‽ How is that possible? Basically, coders focus on applying the same operation, row by row, over a data frame --- without toiling over the syntax of For loops.

[^1]: <https://en.wikipedia.org/wiki/Functional_programming>

[^2]: [https://en.wikipedia.org/wiki/For_loop](https://en.wikipedia.org/wiki/For_loophttps://en.wikipedia.org/wiki/For_loop)

If you're new to R, you may not realize that you've been using functions throughout this workshop series. All the packages we introduced (dplyr, ggplot2, skimr, and even base-R, etc.) are made up collections of functions. For example, {dplyr} gives us the `filter()` and the `select()` functions. At the end of the [section on regression](regression.html#example-of-iterative-modeling-with-nested-categories-of-data), we saw an example of using the `nest()` function so our data frame is set for the repetition of our regression (the `lm()` function) row-by-row. We did this without ever writing a FOR loop.

## Definition

::: callout-note
## Functions

Blocks of code that can be invoked by the coder to perform fundamental steps.
:::

Functions can accept inputs and yield outputs. If we put a string of functions together we have a more complex function. We may want to apply this complex function to some group of data till we reach the end. In R the basic syntax of a function looks like this

::: column-margin
<iframe width="300" height="200" src="https://www.youtube.com/embed/EV3pOavW95g" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" allowfullscreen>

</iframe>
:::

```{{r}}

# Define function to add two numbers
add_numbers <- function(x, y) { x + y }

```

Then we can call that function like this

```{{r}}
add_nubmers(8, 10)
```

The console will return a response, or we can store that response in an object. Or we can combine that function with the mutate function and iterate over all the rows in a data frame

```{r}
#| warning: false
#| message: false

library(tidyverse)

add_numbers <- function(x, y) {
  x + y
}

starwars |> 
  mutate(my_product = add_numbers(mass, height), .after = mass)
```

While the custom function, `add_numbers`, didn't save us much time. We can create more sophisticated functions. For example, we can write a function that plots x and y.

The function, below, uses [*indirection syntax*](https://dplyr.tidyverse.org/articles/programming.html#indirection) (i.e. double-braces) because there's a difference between an *environment* variable and a *data* variable. An explanation of indirection is found in the enthusiastically recommended [*Programming with dplyr*](https://dplyr.tidyverse.org/articles/programming.html)vignette. Essentially, an **environment variable** shows up in the RStudio environment tab. While a **data** variable is the name of a variable *within* an environment variable. For example, in a data frame, a column header variable name is a data variable. Data variables have to be called with the [*embrace* operator `{{`](https://rlang.r-lib.org/reference/embrace-operator.html)so that they data can be referenced indirectly. This indirection overcomes a [data masking feature](https://rlang.r-lib.org/reference/topic-data-mask.html) which allows for more efficient coding in many cases.\
\
Essentially, ease of coding is emphasised with data masking. This data masking makes it easier to learn R. As a beginner there are fewer syntactical barriers when coding in a tidyverse context. The paradox of this early simplicity comes at the cost of more complex syntax when composing complex functions. This syntactical complexity occurs because data-masked variables need to be referred to indirectly.

```{r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Plot 1"
#|   - "Plot 2"

make_scatterplot <- function(my_df, my_x, my_y, my_color) {
  my_df |>
    drop_na() |>
    ggplot(aes(x = {{my_x}}, y = {{my_y}})) +
    geom_point(aes(color = {{my_color}})) +
    geom_smooth(method = lm, se = FALSE, formula = y ~ x)
}

make_scatterplot(starwars, height, mass, gender)

make_scatterplot(mpg, displ, hwy, class)

```

::: callout-tip
## Rule of thumb

If you have to write the same code three times or more, write a function. Among the advantages, this prevents typographical errors. Additionally, writing functions can prevent mistakes in coding because there will be fewer places to update.
:::

## Can it be simpler?

In a word, yes, but there may be limitations. In the [section on pivoting data](longer_wider.html#why-pivot-data), we saw how pivoting data to the tall format made it easier to generate multiple {ggplot2} plots. We combine pivoting with the `facet_wrap()` function to visualize multiple plots with minimal coding. While we didn't write a special custom function, we did use the functional programming approach to iterate over rows of a data frame. Without pivoting the data and faceting the plots, this code might have taken ten-times as much code, most of it repetitive and all of it susceptible to typing mistakes.

```{r}
#| fig-height: 4
#| fig-width: 10

inc_levels = c("Don't know/refused",
               "<$10k", "$10-20k", "$20-30k", "$30-40k",
               "$40-50k", "$50-75k", "$75-100k", "$100-150k",
               ">150k")

relig_income %>%
  pivot_longer(-religion, names_to = "income", values_to = "count") %>% 
  mutate(religion = fct_lump_n(religion, 4, w = count)) %>% 
  mutate(income = fct_relevel(income, inc_levels)) %>% 
  summarise(sumcount = sum(count), .by = c(religion, income)) %>% 
  ggplot(aes(fct_reorder(religion, sumcount), 
             sumcount)) + 
  geom_col(fill = "grey80", show.legend = FALSE) +
  geom_col(data = . %>% filter(income == "$40-50k"),
           fill = "firebrick") +
  geom_col(data = . %>% filter(income == ">150k"),
           fill = "forestgreen") +
  coord_flip() +
  facet_wrap(vars(income), nrow = 2) 
```

## Mapping over a data frame

But how do we apply functions, row-by-row, over a data frame without using FOR loops? Move to the section on [iteration with {purrr}](purrr.html)
