---
title: "GIS"
output: html_document
#  html_notebook:
#    css: duke.css
---

We offer a separate workshop on mapping in R.  Please see our [*Mapping with R* guide](https://map-rfun.library.duke.edu/) for more information.  

```{r loadPackages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(mapview)

```


```{r loadData, message=FALSE, warning=FALSE}
starbucks <- read_csv("https://github.com/data-and-visualization/Intro2R/raw/master/data/All_Starbucks_Locations_in_the_US_-_Map.csv")
```


## GIS Mapping

Introducing the [`mapview` package](https://r-spatial.github.io/mapview/index.html).  The demonstration code below will use some latitude and longitude data to generate an interactive map.  The basemap can be adjusted, and the final map can also be exported as an HTML file or a static file. This example is adapted from an article by Machlis.^[[5 Visualizations in 5 Minutes](http://www.computerworld.com/article/2893271/business-intelligence/5-data-visualizations-in-5-minutes-each-in-5-lines-or-less-of-r.html). ComputerWorld.com by Sharon Machlis]

### Filter Data to NC
```{r filter-dataset}
starbucksNC <- starbucks  %>% 
  filter(State == "NC")
```

### Make the Map

In this example, set the map projection to a common projection standard, WGS84, via the argument `crs = 4326`.)

```{r map1}
mapview(starbucksNC, xcol = "Longitude", ycol = "Latitude", crs = 4269, grid = FALSE)
```


### Alternative: Transform data to Spatial object
```{r map2}
starbucksNC2 <- st_as_sf(starbucksNC, coords = c("Longitude", "Latitude"),  crs=4326)
```


### Make the Map
```{r makemap}
mapview(starbucksNC2)
```

## Save the Map

You can save a map as an interactive HTML page or a static `.png, .pdf, or .jpeg` file with the [`mapshot` function](https://r-spatial.github.io/mapview/reference/mapshot.html).  For example:

``` r
starNCmap <- mapview(starbucksNC, xcol = "Longitude", ycol = "Latitude", crs = 4269, grid = FALSE)
```

- `mapshot(starNCmap, file = "map.png")`
- `mapshot(starNCmap, url = "map.html")`

Learn more about mapping in R

- [Our workshop on ggplot2](https://rfun.library.duke.edu/portfolio/mapping_workshop/). We offer free video streaming with shareable materials as well.  
- [Documentation for mapview](https://r-spatial.github.io/mapview/) 
- [Documentation for sf](https://r-spatial.github.io/sf/)


