---
title: "Interactivity: Quarto & Observable JS"

date-modified: 'today'
date-format: long

license: CC BY-NC 
bibliography: references.bib

format: 
  html:
    df-print: paged
    toc: true
    echo: false
    warning: false
    messages: false
    code-summary: "Show the code"
    code-tools: true
---

```{r}
library(tidyverse)
library(haven)
satisfaction_df <- read_sav("data/student_satisfaction_test-data_from_qualtrics.sav") |> 
  janitor::clean_names() |> 
  mutate(across(is.labelled, \(x) as_factor(x), .names = "{.col}_label")) |> 
  select(response_id, ends_with("_label")) |> 
  pivot_longer(-response_id, names_to = "my_question", values_to = "response") |> 
  mutate(rnumber = as.numeric(response))

trade_df <- read_csv("data/data.csv")
trade_df <- trade_df |> 
  filter(str_detect(business, "(Motor|beverage|Electronics)")) |> 
  mutate(sales = as.integer(sales))
```

```{r}
ojs_define(trade_df_ojs = trade_df)
ojs_define(my_df_tall_ojs = satisfaction_df)
```

This is a very basic example of displaying an interactive visualization using ObservableJS. Use the drop-down box to test how the plots and tables are reactive to user input.

```{ojs}
//| panel: input
//| layout-ncol: 1

viewof questions = 
Inputs.select(
  ["q1_label", "q2_label", "q3_label", "q4_label", "q5_label",
   "q6_label", "q7_label", "q8_label", "q9_label", "q10_label",
   "q12_label", "q13_label", "q14_label", "q15_nps_group_label"], 
  {value: "q3_label", label: "Select a question"})
  
// viewof questions = Inputs.select(d3.group(my_survey, d => d.my_question))

```

::: panel-tabset
## Plot

```{ojs}
//| code-fold: true
//| echo: fenced

my_survey_filtered_too = my_survey
  .filter(aq.escape(d => d.my_question == questions))
  .groupby('my_question', 'response', 'rnumber')
  .count()
  .orderby('my_question', 'rnumber')
  
Plot.plot({
  marginRight: 140,
  y: {
    axis: null,
    domain: d3.sort(my_survey_filtered_too, d => -d.rnumber).map(d => d.response).reverse()
  },
  x: {
    grid: true
  },
 marks: [
    Plot.barX(my_survey_filtered_too, {y: "response", x: "count"}),
    Plot.text(my_survey_filtered_too, {
      x: "count",
      y: "response",
      text: "response",
      textAnchor: "start",
      dx: 6
    })
  ]
})
```

## Data

```{ojs}
my_survey_filtered_too.view()
```
:::

::: {.callout-tip collapse="true"}
## Workflow: How'd we do that?

The above example uses `{r}` code-chunks and the `{haven}` package to import labeled Qualtrics data. The SPSS data file is a .*sav* file with raw data, variable and value labels ingested via `read_sav`, wrangled with `{dplyr}`, and then visualized with client-side interactive ObservableJS `{ojs}` code-chunks. Using the RStudio IDE and Quarto, it's easy to combine R and Observable code so that we can render this interactive page. (See the `</> Code` icon at the top of this page.)

Learn more below...
:::

## Transpose to arquero

Behind the scenes, when rendering this report, I use `{r}` code-chunks to import and wrangle the data. I'm importing and wrangling my data with a quarto-friendly code language. Then I *transpose* my data frame into an OJS object.

```{ojs}
//| echo: fenced
my_data = aq.from(transpose(trade_df_ojs))
my_survey = aq.from(transpose(my_df_tall_ojs))
```

## Simplified example

### basic

The next code-chunk shows a very simple example of ObservableJS `{Plot}` code for visualization. This example *is not interactive* but it demonstrates the remarkable similarity to `{ggplot2}` syntax.

plot

```{ojs}
//| echo: fenced
Plot.plot({
  marks: [
    Plot.ruleY([0]),
    Plot.lineY(my_data_wrangled, {
      x: "date",
      y: "sales",
      stroke: "business"
    }),
  ]
})
```

::: callout-note
## ↥ Look up!

Notice the crispness and clarity of the simple non-interactive plot above. [Observable Plot](https://observablehq.com/@observablehq/plot) shows well on the web!
:::

## Arquero data wrangling

::: gray
**Derive date**

Aside from the `{dplyr}` wrangling that I did in the background of this code notebook, I still needed to parse the date fields. Arquero to the rescue.

**Arquero**! Now what is that? *Observable JavaScript* is its own coding language and we can minimize the learning curve by keeping our data wrangling within our favorite coding language R. This way we can focus on using the [{Plot}](#0) grammar and syntax that is very similar to `{ggplot2}` -- see the above example. However, [Observable JS](#0) is a complete and rich language of its own. So, just like [{`shiny`}](#0), we want to learn just enough to get the job done. That said, there is a case where a little more data wrangling is optimal. This is where [`{arquero}`](#0) comes in. Arquero is used for data wrangling and is easy to learn because it was inspired by `{dplyr}`. *Show the code* for the first plot on this page and you'll see the similarity to `{dplyr}`.

Below we can coerce a data-type for the `my_data$date` vector by transforming a *character* data-type to a *date* data-type using Arquero's `.derive()` function. Arquero's `.derive()` function is similar to `dplyr::mutate()`
:::

```{ojs}
//| echo: fenced
my_data_wrangled = my_data
  .derive({  date: d =>  op.parse_date(d.date) })
```

On this page, we use both `{r}` and `{ojs}` code chunks within one Quarto notebook. This shows how to present useful interactivity in a multilingual coding context.

```{ojs}
import { aq, op } from '@uwdata/arquero'
// aq = import('https://esm.run/arquero') 
// op = import('https://esm.run/arquero')
```

## Examples and documentation

See a [very cool demo at the Quarto-Observable](https://quarto.org/docs/interactive/#observable-js) introduction page. Learn more about this exciting option at the [*Quarto-Observable JS Guide*](https://quarto.org/docs/interactive/ojs/).
