---
title: "Iteration with {purrr}"

date-modified: 'today'
date-format: long

license: CC BY-NC
bibliography: references.bib
---

::: {.callout-warning collapse="true"}
## Review the functions page!

If you have not already read the [functions page](functions.html), do so before reading about {purrr}. The purrr package is about applying functions repetitively. You should have a good idea about what a function is and how to write a custom function.
:::

From the [functions page](functions.html) we learned scant definitions of *functional programming* and *functions*. We also learned about the special conditions of programming with {dplyr}, which demands a working understanding of *environment* and *data* variables.[^1] Now we want to apply our functions, row-by-row, to a data frame. We'll use the [purrr package](https://purrr.tidyverse.org/).

[^1]: Links and footnotes on the [functions page](functions.html) will lead to more detailed information on those topics.

::: callout-tip
## Rule of thumb

Never feel bad for using a FOR loop.
:::

Remember that in functional programming we're iterating, or recursing, without using FOR loops. For example, in the [regression page](regression.html#example-of-iterative-modeling-with-nested-categories-of-data), we saw an example of *nesting data frames* by category (i.e. by `gender`). After nesting, we have two subsetted data frames that are embeded within a parent-data-frame. The embedded data frames are contained within a *list-column*.

### `nest()`

We can subset by gender, creating two new subset data frames that we'll put in a new column with the variable name: *data*. This subsetting is accomplished with the `nest()` function.

::: column-margin
<iframe width="300" height="200" src="https://www.youtube.com/embed/kZf11zbVpr0" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" allowfullscreen>

</iframe>
:::

```{r}
#| warning: false
#| message: false

library(moderndive)
library(broom)
library(tidyverse)

evals |> 
  janitor::clean_names() |> 
  nest(data = -gender)
```

## `map()`

Using {purrr} we can apply a function --- such as a linear model regression `lm()` --- over each row of the parent data frame. To do this, we use the `map_` class of functions. I say *class* *of functions*, because `purrr::map_` allows us to define the data-type[^2] returned by the mapped function. For example, sometimes we want to return a *character* data-type (`map_chr`), sometimes an *integer* (`map_int`), sometimes a data frame (`map_df`), etc. In the case of a linear model, which is a *list* data-type, we'll use the generic `map()` function to apply an ***anonymous** function*.

[^2]: A fuller explanation of data types can be found in *R for Data Science* [@wickham2023]

::: column-margin
<iframe width="300" height="200" src="https://www.youtube.com/embed/QgasjZGhWlk" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" allowfullscreen>

</iframe>
:::

::: callout-note
## Anonymous functions

Examples of an anonymous function:

`\(x) x + 1`

`function(x) x + 1`

Unlike the `add_numbers()` function we composed on the [functions page](functions.html#definition), anonymous functions have no object name. Hence they are **anonymous**. Anonymous functions, sometimes called lambda functions, are a convenience for coders using functional programming languages.
:::

```{r}
#| warning: false
#| message: false

my_subset_df <- evals |> 
  janitor::clean_names() |> 
  nest(my_data = -gender) |> 
  mutate(my_fit_bty = map(my_data, function(my_data) 
                                lm(score ~ bty_avg, data = my_data)
                         )
        )
my_subset_df
```

### Look at the fitted object

When we look at the resulting `my_fit_bty` data variable, we see the kind of output we get from `lm()`

```{r}
my_subset_df$my_fit_bty
```

Earlier, we learned to use the `broom::tidy()` function to transform a fitted model --- contained as a *list* data-type --- into a data frame.[^3] For example, we can easily `tidy` the first model which coerces the list data-type into a data frame.

[^3]: Having our regression model wrapped as a data frame means we can use other {dplyr} functions to more easily manipulate our model output.

```{r}
tidy(my_subset_df$my_fit_bty[[1]])
```

### Named functions

If we wanted to tidy all the models --- not only the fist model (above) --- then we use the `map` function (again.) This time we **map** a named function, i.e. `broom::tidy()`. We do this row-by-row with `map()`, i.e. **without** using a FOR loop.

```{r}
my_subset_df |> 
  mutate(my_fit_bty_tidy = map(my_fit_bty, tidy))
```

A difference between `my_fit_bty` and `my_fit_bty_tidy` is that `my_fit_bty_tidy` is the model output for the former is a *list*, the later is a *data frame*. Therefore, the parent data frame --- `my_subset_df`--- has a data variable: `my_fit_bty_tidy`, aka `my_subset_df$my_fit_bty_tidy`. my_fit_bty_tidy is a list-column of nested data frames, just like `my_subset_df$data` --- which we nested in the first code-chunk of this page as `evals$data`.

To un-nest the data frames, use the `unnest()` function.

```{r}
my_subset_df |> 
  mutate(my_fit_bty_tidy = map(my_fit_bty, tidy)) |> 
  unnest(my_fit_bty_tidy)
```

And now we have fitted model data, contained as tidy-data, within a data frame, which we can manipulate further with other tidyverse functions. For example, using dplyr functions it's easy to make a pipeline to look at the p-values of `bty_avg`, by `gender`.

```{r}
my_subset_df |> 
  mutate(my_fit_bty_tidy = map(my_fit_bty, tidy)) |> 
  unnest(my_fit_bty_tidy) |> 
  select(gender, term, p.value) |> 
  filter(term == "bty_avg")
```

When we combine linear modeling with other broom functions such as `glance()` and `augment()`, then we can build on our analysis and data manipulation.

## Anonymous functions

Anonymous functions have no name. Unlike a **named** function (e.g. `broom::tidy` above, or `make_scatterplot` from the [functions page](functions.html)). Below we **map** an anonymous function within `map()`. The first argument to map is a *list* **or** *data frame*. The next argument can be a named function **or** an anonymous function. In the example below, the anonymous function produces a scatter plot with a regression line.

```{r}
my_subset_df_with_plots <- my_subset_df |> 
  mutate(my_plot = map(my_data, function(my_data) 
    my_data |> 
      ggplot(aes(bty_avg, score)) + 
      geom_point() +
      geom_smooth(method = lm, se = FALSE, formula = y ~ x)
                      )
         )
my_subset_df_with_plots
```

And now we can pull those plots

```{r}
#| eval: false
#| echo: true
my_subset_df_with_plots |> 
  pull(my_plot)
  
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-cap: 
#|   - "Female"
#|   - "Male"


my_subset_df_with_plots$my_plot[[1]]

my_subset_df_with_plots$my_plot[[2]]
```

## `map2()` & `pmap()`

If you have more than one argument to map, you can use functions such as `map2()` or `pmap()`. An example of `map2()` can be found on the [regression page](regression.html#example-of-iterative-modeling-with-nested-categories-of-data).

```{{r}} 
# Example of mapping a anonymous function with two variables.

map2(my_df, my_plot, gender, function(x, y) { x + labs(title = str_to_title(y)) } )

# x refers to the second argument of map2, i.e. `my_plot`

# y refers to the third argument, i.e. `gender`

```

 
