---
title: "Example EDA"
author: "John Little"

date-modified: 'today'
date-format: long

license: CC BY-NC
bibliography: references.bib
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(skimr)
```

Basic steps taken in this rough exploration of data....

-   import data
-   wrangle data
-   join data
-   visualize data

A brief discussion of packages which proport to perform EDA for you can be found in the [Get Started section of this site](eda.html "selective summary of eda packages").

Note: This page is not intended to teach formal EDA. What happens on this page is merely a brutal re-enactment of some informal explorations that a person might take as they familiarize themselves with new data. If you're like most people, you might want to [skip to the visualizations](#visualize-wrangle-and-summarize). Meanwhile, the sections and code-chunks preceding visualization are valuable and at least worth a glance,

## Import data

The data come from a [TidyTuesday](https://github.com/rfordatascience/tidytuesday), a weekly social learning project dedicated to gaining practical experience with R and data science. In this case the TidyTuesday data are based on [International Trail Running Association (ITRA)](https://itra.run/Races/FindRaceResults) data but inspired by Benjamin Nowak, . We will use the [TidyTuesday data that are on GitHub](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-10-26). Nowak's data are [also available on GitHub](https://github.com/BjnNowak/UltraTrailRunning).

```{r}
#| message: false
#| warning: false
race_df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-26/race.csv")
rank_df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-26/ultra_rankings.csv")
```

```{r}
glimpse(race_df)
```

```{r}
glimpse(rank_df)
```

## EDA with skimr

```{r}
skim(race_df)
```

Read more about automagic [EAD packages](eda.html).

## Freewheelin' EDA

```{r}
race_df |> 
  count(country, sort = TRUE) |> 
  filter(str_detect(country, regex("Ke", ignore_case = TRUE)))
```

```{r}
race_df |> 
  filter(country == "Turkey")
```

```{r}
race_df |> 
  count(participation, sort = TRUE)
```

```{r}
race_df |> 
  count(participants, sort = TRUE)
```

```{r}
skim(rank_df)
```

```{r}
rank_df |> 
  filter(str_detect(nationality, regex("ken", ignore_case = TRUE)))
```

```{r}
rank_df |> 
  arrange(rank)

rank_df |> 
  count(rank, sort = TRUE)
rank_df |> 
  drop_na(rank) |> 
  count(rank, gender, age, sort = TRUE)

race_df |> 
  count(distance, sort = TRUE)
```

```{r}
rank_df |> 
  filter(race_year_id == 41449)

race_df |> 
  filter(race_year_id == 41449)

race_df |> 
  filter(distance == 161)
```

```{r}
race_df
```

```{r}
race_df |> 
  count(race, city, sort = TRUE)

race_df |> 
  filter(race == "Centurion North Downs Way 100")
```

```{r}
race_df |> 
  filter(race_year_id == 68140)

race_df |> 
  filter(race == "Millstone 100")

race_df |> 
  filter(event == "Peak District Ultras")

race_df |> 
  count(race, sort = TRUE)  

race_df |> 
  count(city, sort = TRUE)  

race_df |> 
  count(event, sort = TRUE)  
```

```{r}
race_df |> 
  filter(event == "Burning River Endurance Run")
```

## Visualize, wrangle, and summarize

Here I'm using [this *State of Ultra Running* report](https://runrepeat.com/state-of-ultra-running) as a model to demonstrate **some** of the capabilities of R / Tidyverse

### join datasets

#### Join, Assign, and Pipe

In this case I want to **join** the two data frames `rank_df` and `race_df` using the `left_join()` function.

I can **assign** the output of a "data pipe" (i.e. data sentence) to use in subsequent code-chunks. A common R / Tidyverse assignment operator is the `<-` characters. You can read this as "gets value from".

Additionally, I'm using a **pipe** operator (`|>`) as a conjunction to connect functions. In this way I can form a *data sentence*. Many people call the data sentence a *data pipe*, or just a *pipe*. You may see another common pipe operator: `%>%`. `\>` and `%>%` are synonymous.

using `dplyr::left_join()` I combine the two data sets and then use {`ggplot2`} to create a line graph of participants by year.

```{r}
my_df_joined <- rank_df |> 
  left_join(race_df, by = "race_year_id") |>
  mutate(my_year = lubridate::year(date))
```

### Viz participants

Let' make a quick line plot showing how many people participate in races each year. Here we have a `date` field this is also a date data-type. Data types are important and in this example using a data data-type means {`ggplot2`} will simplify our x-axis labels.

Here we use the {`lubridate`} package to help manage my date data-types. We also use {`ggplot2`} to generate a line graph as a *time series* via the {`ggplot2`} package and a `geom_line()` *layer*. Note that {`ggplot2`} uses the '`+`' as the conjunction or *pipe*.

```{r}
rank_df |> 
  left_join(race_df |> select(race_year_id, date), by = "race_year_id") |>
  mutate(my_year = lubridate::year(date)) |> 
  count(my_year, sort = TRUE) |> 
  ggplot(aes(my_year, n)) +
  geom_line()
```

### by distance

Here I use `count()` in different ways to see what I can see. I comment out each attempt before settling on summarizing a table of total country participants by year.

```{r}
my_df_joined |> 
  mutate(participation = str_to_lower(participation)) |> 
  # count(participation, sort = TRUE)
  # count(city) |> 
  # count(race) |> 
  count(my_year, country, sort = TRUE)

my_df_joined |> 
  mutate(participation = str_to_lower(participation)) |> 
  count(my_year, country, sort = TRUE) |> 
  drop_na(country) |> 
  mutate(country = fct_lump_prop(country, prop = .03)) |> 
  ggplot(aes(my_year, n)) +
  geom_line(aes(color = country))

```

### by country

I used `fct_lump_prop()` in the previous code-chunk to lump the `country` variable into categories by frequency. Here we refine the categories into specific **levels**. We are still *mutating* the `country` variable as a categorical factor; this time using the `fct_other()` function of {`forcats`} with some pre-defined levels (see the `my_levels` vector in the code-chunk below).

```{r}
my_levels <- c("United States", "France", "United Kingdom", "Spain")

my_df_joined |> 
  mutate(country = fct_other(country, keep = my_levels)) |> 
  count(my_year, country, sort = TRUE) |> 
  drop_na(country) |> 
  ggplot(aes(my_year, n, color = country)) +
  geom_line() +
  geom_point() +
  scale_color_brewer(palette = "Dark2") 
```

### Country race-host

```{r}
my_df_joined |> 
  drop_na(country) |> 
  mutate(country = fct_lump_n(country, n = 7)) |>
  count(country, sort = TRUE) |> 
  ggplot(aes(x = n, y = fct_reorder(country, n))) +
  geom_col()
```

### Nationality of runner

```{r}
my_df_joined |> 
  mutate(nationality = fct_lump_n(nationality, n = 7)) |> 
  count(nationality, sort = TRUE) |> 
  ggplot(aes(n, fct_reorder(nationality, n))) +
  geom_col()
```

### Unique participants

```{r}
my_df_joined |> 
  distinct(my_year, runner) |> 
  count(my_year) |> 
  ggplot(aes(my_year, n)) +
  geom_line()
```

### Participant frequency separated by gender

Note the use of `count`, `if_else`, `as.character`, and `group_by` to transform the data for visualizing. Meanwhile, the visual bar graph is a proportional graph with the y-axis label by percentage. We do this by manipulating the plot *scales*. Scales are also used to choose colors from a predefined palette (i.e. "Dark2".) Findally, we facet the plot by gender (See `facet_wrap()`).

```{r}
my_df_joined |> 
  count(my_year, gender, runner, sort = TRUE) |> 
  mutate(n_category = if_else(n >= 5, "more", as.character(n))) |> 
  group_by(my_year) |> 
  mutate(total_races = sum(n)) |> 
  ungroup() |> 
  ggplot(aes(my_year, total_races))  +
  geom_col(aes(fill = fct_rev(n_category)), position = "fill") +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(vars(gender)) 
```

### Pace per mile

We want to calculate a value for each runner's pace (i.e. `minute_miles`). We have to create and convert a character data-type of the `time` variable into a numeric floating point (or `dbl`) data-type so that we can calculate pace (i.e. race-minutes divided by distance.) These data transformations required a lot of manipulation as I was thinking through my goal. I could optimized this code, perhaps. However it works and I've got other things to do. Do I care if the CPU works extra hard? No, not in this case.

```{r, dev='svg'}
my_df_joined |> 
  mutate(time_hms = str_remove_all(time, "[HMS]"), .after = time) |> 
  mutate(time_hms = str_replace_all(time_hms, "\\s", ":")) |> 
  separate(time_hms, into = c("h", "m", "s"), sep = ":") |> 
  mutate(bigminutes = (
    (as.numeric(h) * 60) + as.numeric(m) + (as.numeric(s) * .75)
    ), .before = h) |> 
  mutate(pace = bigminutes / distance, .before = bigminutes) |> 
  drop_na(pace, distance, my_year) |> 
  filter(distance > 0,
         pace > 0) |> 
  group_by(my_year, gender) |>  
  summarise(avg_pace = mean(pace), max_pace = max(pace), min_pace = min(pace)) |> 
  pivot_longer(-c(my_year, gender), names_to = "pace_type")  |> 
  separate(value, into = c("m", "s"), remove = FALSE) |> 
  mutate(h = "00", .before = m) |> 
  mutate(m = str_pad(as.numeric(m), width = 2, pad = "0")) |> 
  mutate(s = str_pad(round(as.numeric(str_c("0.",s)) * 60), width = 2, pad = "0")) |> 
  unite(minute_miles, h:s, sep = ":") |> 
  mutate(minute_miles = hms::as_hms(minute_miles)) |> 
  # drop_na(gender) |> 
  ggplot(aes(my_year, minute_miles)) +
  geom_line(aes(color = pace_type), size = 1) +
  scale_color_brewer(palette = "Dark2") +
  theme_classic() +
  facet_wrap(vars(gender))
```

### Age trends

In this code-chunk we use a {`ggplot2`} function, `cut_width()`, to generate rough categories by age. `dplyr::case_when()` is a more thorough and sophisticated way to make some cuts in my data, but `ggplot2::cut_width()` works well for a quick visualization.

Note the use of labels, scales, themes, and guides in the last visualization. A good plot will need refinement with some or all of these functions.

```{r}
my_df_joined |> 
  mutate(age_cut = cut_width(age, width = 10, boundary = 0), .after = age) |> 
  count(age_cut, gender, sort = TRUE)

my_df_joined |> 
  filter(age < 80) |> 
  drop_na(gender) |> 
  ggplot(aes(y = cut_width(age, width = 10, boundary = 0))) +
  geom_bar(aes(fill = gender)) +
  facet_wrap(vars(gender))

my_df_joined |> 
  filter(age < 70, age >= 20) |>
  drop_na(gender) |> 
  ggplot(aes(my_year)) +
  geom_bar(aes(fill = fct_rev(cut_width(age, width = 10, boundary = 0))), position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Dark2") +
  labs(fill = "Age", title = "Age distribution of ultra runners",
       caption = "Source: ITRA > Benjamin Nowak > Tidy Tuesday",
       x = NULL, y = NULL) +
  theme_classic() +
  theme(legend.position = "top", plot.title.position = "plot") + 
  guides(fill = guide_legend(reverse = TRUE))

```

 
