---
title: "Visualization"
output: html_document
#  html_notebook:
#    css: duke.css
---

When in the data exploratory phase, you can quickly generate very basic visualizations with minimal effort.   

For this introductory section on visualization we use the `ggplot2` package.  We offer a [separate workshop on ggplot2](https://rfun.library.duke.edu/portfolio/ggplot_workshop/).    [Documentation for ggplot2](http://ggplot2.tidyverse.org/) is readily available.  

You can learn more about ggplot with the following free resources

- Interactive Learning:  [Visualize Data Primers](https://rstudio.cloud/learn/primers/3)
- [Data Visualization](https://r4ds.had.co.nz/data-visualisation.html) Chapter in *R for Data Science* by Wickham & Grolemund.  See also the Chapter, "Graphics for Communication".

```{r loadPackages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Basic Grammar

The grammar template for a ggplot2 plot can be seen below. The `ggplot` function takes a dataframe as the first primary input.  Variables can be defined within the aesthetics argument, and finally [various geometries](https://ggplot2.tidyverse.org/reference/index.html#section-layer-geoms) are defined. Unlike all other Tidyverse packages, ggplot2 **uses a plus operator `+` for a pipe**.  This is **different from `%>%`** but ultimately not that hard to implement. 

```r
ggplot(data = starwars, aes(x = height, y = mass)) +
  geom_----()
```

**Some Example Visualizations and Geometries**

## Scatter Plot

Use [`geom_point`](https://ggplot2.tidyverse.org/reference/geom_point.html)

```{r geompoint, warning=FALSE}
ggplot(data = starwars, aes(x = mass, y = height)) +
  geom_point()
```


## Bar Plot

[`geom_bar`](https://ggplot2.tidyverse.org/reference/geom_bar.html) - counts the cases and presents the bar heights are proportional to the number of cases.

[`geom_col`](https://ggplot2.tidyverse.org/reference/geom_bar.html) - bar heights represent the values in the data

```{r}
ggplot(data = starwars) +
  geom_bar(aes(x = fct_infreq(eye_color)))
```

A more refine example that also demonstrates how to manipulate labels and colors
```{r}
ggplot(starwars) +
  geom_bar(aes(x = fct_rev(fct_lump(fct_infreq(eye_color), n = 6))), 
           fill = "grey") + 
  coord_flip() +
  geom_bar(data = starwars %>% filter(eye_color == "orange"),
           aes(eye_color), fill = "orange") +
  labs(x = "Eye Color", y = "",
       title = "Eye Color of Starwars Characters",
       subtitle = "Oranges eyes are uncommon.",
       caption = "Data Source: dplyr::starwars")

```


## Histogram

[`geom_histogram`](https://ggplot2.tidyverse.org/reference/geom_histogram.html)
```{r hist, message=FALSE, warning=FALSE}
ggplot(data = starwars, aes(x = birth_year)) + 
geom_histogram()
```

Bins can be adjusted with `bindwidth` or `bins`.  See the documentation for arguments
```{r binwidth, warning=FALSE}
ggplot(data = starwars, aes(x = birth_year)) + 
geom_histogram(binwidth = 10)
```

## BoxPlot

[`geom_boxplot`](https://ggplot2.tidyverse.org/reference/geom_boxplot.html)
```{r boxplot, message=FALSE, warning=FALSE}
starwars %>% 
  filter(species == "Human" | species == "Droid" | species == "Gungan") %>% 
  ggplot(aes(x = species, y = mass)) +
    geom_boxplot()
```

## Grouping + Legends

Variables can be mapped to exhibit different aesthetics.  In this case we map the `species` variable to the `color` aesthetic and `birth_year` variable to `size` aesthetic.  We also account for overplotting by adjusting the opacity as a fixed argument (outside the `aes()`) with the `alpha` argument.  Variables can be mapped inside the aesthetics function while other fixed characteristics are mapped outside the aesthetic function but inside the geom_.   **Legends** are added automatically by ggplot2.  

```{r grouping}
starwars %>% 
  drop_na(mass, height, birth_year) %>% 
  filter(mass < 400,
         species == "Human" | 
           species == "Droid" | 
           species == "Gungan" |
           species == "Mirialan" |
           species == "Ewok" |
           species == "Wookiee") %>% 
  ggplot(aes(x = mass, y = height)) +
    geom_point(aes(color = species, size = birth_year), alpha = 0.7)
```



Learn more about ggplot2.

- [Our workshop on ggplot2](https://rfun.library.duke.edu/portfolio/ggplot_workshop/) We offer free video streaming with shareable materials as well.  
- Interactive Learning:  [Visualize Data Primers](https://rstudio.cloud/learn/primers/3)
- [Documentation for ggplot2](http://ggplot2.tidyverse.org/) is readily available. 
- [Data Visualization](https://r4ds.had.co.nz/data-visualisation.html) Chapter in *R for Data Science* by Wickham & Grolemund.  See also the Chapter, "Graphics for Communication"