---
title: "Visualization"
output: html_document
#  html_notebook:
#    css: duke.css
---

When in the data exploratory phase, you can quickly generate very basic visualizations with minimal effort.   

For this introductory section on visualization we use the `ggvis` package because it works seamlessly with pipes (i.e. `%>%`).  You can do a lot with [ggvis](http://ggvis.rstudio.com/).  You can do even more with [ggplot2](http://ggplot2.tidyverse.org/).  We offer a [separate workshop on ggplot2](https://rfun.library.duke.edu/2017/09/26/visualization-in-r-using-ggplot2-workshop-fall-2017/).

```{r loadPackages, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggvis)
```

## Wrangle Data

Create a data frame without Jabba
```{r not_jabba}
not_jabba <- starwars %>% 
  filter(!is.na(mass),
         !is.na(height),
         mass < 400)
```

Subset the Species list to six species.

```{r ggvis_prepare_bargraph}

starwars_species_small <- starwars %>% 
  filter(!is.na(mass),
         !is.na(height),
         !is.na(birth_year),
         !is.na(species),
         mass < 400,
         species == "Human" | 
           species == "Droid" | 
           species == "Gungan" |
           species == "Mirialan" |
           species == "Ewok" |
           species == "Wookiee")

levels <- starwars_species_small %>%
  group_by(species) %>% 
  count() %>% 
  arrange(-n) #%>% 
  #.[1]

starwars_species_small_as_fctr  <-  starwars_species_small %>% 
  mutate(species = parse_factor(species, levels$species))

```

## Scatter Plot
Linear model and 95% confidence interval

```{r ggvis-scatter}
not_jabba %>% 
  ggvis(~height, ~mass) %>% 
  layer_points() %>% 
  layer_model_predictions(model = "lm", se = TRUE)

```
[More information on ggvis scatterplots](http://ggvis.rstudio.com/cookbook.html#scatterplots) | [line graphs](http://ggvis.rstudio.com/cookbook.html#line-graphs)

## Bar Graph


```{r ggvis_bargraph}

starwars_species_small_as_fctr %>%   
  ggvis(~species) %>% 
  layer_bars()
```
[More information on ggvis bar graphs](http://ggvis.rstudio.com/cookbook.html#bar-graphs)

## Histogram

```{r ggvis-hist, message=FALSE}
starwars %>% 
  ggvis(~birth_year) %>% 
  layer_histograms()
```
[More information on ggvis histograms](http://ggvis.rstudio.com/cookbook.html#histograms)

## Grouping data points

```{r}
starwars_species_small_as_fctr %>% 
  ggvis(~height, ~mass) %>% 
  group_by(species) %>% 
  layer_points(~height, ~mass, 
               stroke := "black", 
               fill = ~species,
               size := ~birth_year, 
               opacity := 0.7)
```

&nbsp;

## BoxPlot

```{r, message=FALSE, warning=FALSE}
starwars_species_small_as_fctr %>% 
  filter(species != "Ewok" &
           species != "Wookiee" &
           species != "Gungan") %>% 
  ggvis(~parse_factor(species, species), ~mass) %>% 
  layer_boxplots()
```


[More information on box plots](http://ggvis.rstudio.com/cookbook.html)