---
title: "Easy interactive viz" 
subtitle: "{plotly} and other HTML Widgets"

date-modified: 'today'
date-format: long

license: CC BY-NC
bibliography: references.bib
---

The gg`plotly` function of the {plotly} package [@sievert2020] is an example implementation of [HTML Widgets for R](https://www.htmlwidgets.org/). Beyond {plotly} there are [**many** widgets](http://gallery.htmlwidgets.org/) to choose from. Widgets can be chained together via {[crosstalk](https://rstudio.github.io/crosstalk/)} to support cross-widget interactions such as linked brushing and filtering. The HTML widgets approach is simpler to implement than Shiny and delivers powerful interactive visualizations to your audience.

## Examples

## plotly

In the [previous ggplot2 section on interactivity](viz.html#interactive-plots), we briefly introduced an example interactive bar chart using `ggplotly` to transform a {ggplot2} object. {[plotly](https://plotly.com/ggplot2/)} has its own rich syntax and can do more than make interactive ggplot2 plots. Here's another quick example. Follow the links, above, to more documentation and examples. Try the examples below.

```{r}
#| label: plotly
#| message: false
#| warning: false

library(dplyr)
library(plotly)

my_df <- starwars |> filter(mass < 500) 

plot_ly(my_df, x = my_df$height, y = my_df$mass, 
        text = my_df$name,
        mode = "markers", color = my_df$gender)
  

```

## dygraphs

Another HTML widget is {[dygraph](https://rstudio.github.io/dygraphs/)} for time-series data

```{r}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

## crosstalk and leaflet

Using {[`crosstalk`](https://rstudio.github.io/crosstalk/)} for cross widget interaction (linked brushing and filtering) between two HTML widgets. In this example we link {leaflet} and {DT} widgets

::: column-page
```{r}
#| code-fold: true
#| warning: false
#| message: false

library(crosstalk)
library(leaflet)
library(DT)

# Wrap data frame in SharedData
sd <- SharedData$new(quakes[sample(nrow(quakes), 100),])

# Create a filter input
filter_slider("mag", "Magnitude", sd, column=~mag, step=0.1, width=250)

# Use SharedData like a dataframe with Crosstalk-enabled widgets
bscols(
  leaflet(sd) %>% addTiles() %>% addMarkers(),
  datatable(sd, extensions="Scroller", style="bootstrap", class="compact", width="100%",
    options=list(deferRender=TRUE, scrollY=300, scroller=TRUE))
)
```
:::

## reactable

Or use {[`reactable`](https://glin.github.io/reactable/index.html)} another easy-to-use interactive data tables package.

::: column-page

```{r}
library(reactable)

ggplot2::mpg |> 
  slice_head(n = 20) |> 
  reactable(searchable = TRUE, minRows = 10)

```
:::

## More

See more HTML widgets at the [HTML Widgets for R gallery](http://gallery.htmlwidgets.org/).

In the [next section](interactive.html) we'll introduce a very powerful interactive tool that works with Quarto: ObservableJS.
