---
title: "Visualize Data in R with ggplot2"
subtitle: "Exercises"
abstract: "These exercises are adapted in whole or in part based on the <i>Master the Tidyverse</i> work by Garrett Grolemund at RStudio.  \nCC BY Garrett Grolemund, RStudio ; BY-NC John Little"

date-modified: 'today'
date-format: long

format: 
  html:
    footer: "CC BY 4.0 John R Little"

license: CC BY   
---

These exercises support your learning after watching the [instructional video](https://warpwire.duke.edu/w/80YEAA/)

```{r setup}
#| echo: true
#| warning: false
#| message: false
library(tidyverse)
```

The dataset, `mpg`, is an onboard dataset, part of the `ggplot2` package. You can learn more about it by searching the help tab for "mpg", or type `?mpg` in the console.

```{r}
mpg
```

## Your Turn 1

IN THE code-chunk, **below**, manually type the following code:

ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))

Pay strict attention to spelling, capitalization, and parentheses! Try not to copy and paste, just this once.

```{r}
#| eval: false
```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.

```{r}
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Your Turn 3

Replace this scatter plot with one that draws box plots. Use the cheat sheet. Try your best guess.

```{r}
mpg %>% 
  ggplot() + 
  geom_point(aes(class, hwy))
```

## Your Turn 4

Make a histogram of the `hwy` variable from `mpg`.

```{r}

```

## Your Turn 5

Make a density plot of `hwy` colored by `class`.

```{r}

```

## Your Turn 6

Make a bar chart `hwy` colored by `class`.

```{r}

```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
mpg %>% 
  ggplot() + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))
```

## Your Turn 8

Save the last plot

```{r}
ggsave(_____)
# or right-click the image
```
