---
title: 'Assignment Part 1: Intro2R'
resource_files:
- data/cars.csv
- images/import_data.png
runtime: shiny_prerendered
output: learnr::tutorial
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Loading Data

*Load the file into a new variable called `westUnion`?*

- The data are in a *CSV* file on your local workstation which needs to be loaded into R.  The data file is in a directory relative to your Rmd script: `data/westUnion.csv`.

Write the R code required to load csv data:

```{r read_csv, exercise=TRUE}

```

```{r read_csv-solution}
westUnion <- read_csv("data/westUnion.csv")
# Note...  Code will not run in this tutorial window
# Use an R Notebook to practice this exercise.
```


Hints:

- One option is to use the *Import Dataset* feature of RStudio from the *File* menubar. Code will be generated for you -- which you can run immediately -- or look to the bottom of the dialogue box and copy the code you can run as a code chunk.

## Data Type

```{r datatype}
quiz(
  question("What function will tell you the data type for westUnion?",
    answer("glimpse(westUnion)"),
    answer("tbl_vars(westUnion)"),
    answer("class(westUnion)", correct = TRUE),
    answer("str(westUnion)")
  ),
  
  question("The data type for westUnion is?",
    answer("Data Frame", message = "A modern data frame is known as a Tibble"),
    answer("List"),
    answer("Character"),
    answer("Tibble", correct = TRUE),
    answer("Matrix")
  )
)
```


## Data Structure

Take a look at the structure of the `westUnion` variable.

```{r dataStructure}
quiz(
    question("What is the best command to concisely learn more about the variables, observations, and data types within a Tibble",
             answer("class(westUnion)"),
             answer("glimpse(westUnion)", correct = TRUE),
             answer("tbl_vars(westUnion)"),
             answer("str(westUnion)", message = "This works but is not a conscise report"),
             answer("westUnion", message = "This will work but require a lot of mouse-clicking on your part")
           ),
    
  question("How many observations (rows) are there?",
           answer("59", correct = TRUE),
           answer("4, 4, 6, 8, 7, 8, 9, 4,..."),
           answer("7"),
           answer("None of the Above")
           ),
  
  question("How many variables (columns) are there?",
           answer("59"),
           answer("4 4 6 8 7 8 9 4 9 9 ..."),
           answer("7", correct = TRUE),
           answer("All of the above")
           ),
  
  question("How many of the variables are numeric data?",
           answer("59"),
           answer("7"),
           answer("5"),
           answer("2", correct = TRUE)
           )
  )
```

You have completed the quizzes for Exercise One.  Return to the [Intro2R learning materials](https://rfun.netlify.com/intro2r/), resume at "Data Wrangling".  The quiz below is part of Exercise Two.

## Data Wrangling with dplyr

Pick the best answer for the following questions about dplyr functions

```{r dplyr-quiz}
quiz(
  question("arrange()",
           answer("Sort your data frame rows by an observation", correct = TRUE),
           answer("Make your data frame tidy"),
           answer("Find a perfect match for your data frame"),
           answer("'arrange()' is not a dplyr function"),
           answer("Sort your data frame columns by rows")
    
  ),
  
  question("mutate()",
           answer("Make R into a kitchy horror movie"),
           answer("Transform values in a row of data"),
           answer("Create a new row from other rows"),
           answer("Create a new column", correct = TRUE)
           ),
  
  question("summarize()",
           answer("Tallies a column"),
           answer("Reduces multiple values to a single summary", correct = TRUE),
           answer("Converts an existing table and converts it into a grouped table", message = "'summarize()' is typically used with 'group_by()'"),
           answer("All of the Above")
           ),
  
  question("filter()",
           answer("Removes impurities in the data"),
           answer("Pick variables based on the column name"),
           answer("Adds new variables that are functions of existing variables"),
           answer("Used to limit rows based on values", correct = TRUE)
           ),
  
  question("select()",
           answer("Pick variables based on the column name", correct = TRUE),
           answer("Used to limit rows based on values"),
           answer("Highlights a column variable status for post processing"),
           answer("Changes the ordering of the rows")
           )
)
```

You have completed the quiz for Exercise Two.  Return to the [Intro2R learning materials](https://rfun.netlify.com/intro2r/), resume at "Visualization".  
